What is reported in the news?

### 2017-12-12 phd_slides/phd-workshop.Rmd
final

---
title: "Whats inside the news?"
author: Franziska Löw
date: December 16, 2017
output: 
  revealjs::revealjs_presentation:
    theme: serif
    highlight: haddock
    center: true
    transition: slide
    css: style.css
    self_contained: true
    incremental: true
---

@@

1. **Introduction**

2. **Methodology**

4. **First Results**

5. **Whats next**

@@

- **Does the business model have an effect on the editorial content?**
- **Methodology:**
  1. Estimate a Structural Topic Model
  2. Use posterior distribution to estimate the effect of document metadata.

@@

@@

* Nodes are random variables; arrows indicate dependence
* Plates indicate replicated variables:
    * $N =$ collection of words within a document.
    * $D =$ collection of documents within a corpus.

* Shaded nodes are observed; unshaded nodes are hidden
    * observed: word in a document $w_{d,n}$
    * fixed: mixture components (number of topics $K$ & vocabulary)
    * hidden: mixture proportions (per-document topic proportions $\theta_d$ & word-topic distribution $\beta_k$)

@@

## LDA - joint distribution

$$
\begin{aligned}
p(\beta_{1:K},\theta_{1:D},z_{1:D}, w_{1:D}) \propto
\end{aligned}
$$

$$
\begin{aligned}
\displaystyle \prod_{i=1}^{K}p(\beta_i)\displaystyle \prod_{d=1}^{D}p(\theta_d)(\prod_{n=1}^Np(z_{d,n,}|\theta_d)p(w_{d,n}|\beta_{1:K},z_{d,n}))
\end{aligned}
$$

@@

* **$K$**: choose the number of topics

@@

* **$\theta_d$**: for each document $d$, choose a distribution over topics;

@@

* **$z_{d,n}|\theta_d$**: according to $\theta_d$, assign a topic $z_{d,n}$ for the $n^{th}$ word;

@@

* **$w_{d,n}|z_{d,n},\beta,\theta$**: choose a term from that topic according to $\beta_k$

@@

* **$N$**: repead this process for all $n$ word-positions in the document.

* **$D$**: conduct this process for all $d$ documents in the corpus

@@

## STM Priors

<div style="position:relative; width:600px; height:350px; margin:0 auto;">
  <img class="fragment" src="img/stm5.png" style="position:absolute;top:0;left:0;" />
</div>

@@

$$
\begin{aligned}
p(\eta,z,\kappa,\gamma,\Sigma|w,X,Y) \propto
\displaystyle \prod_{d=1}^{D}(\eta_d|X_d\gamma,\Sigma) (\displaystyle \prod_{n=1}^{D}(z_{n,d}|\theta_d)*(w_n|\beta_{d,k=z_{d,n}}))) * \displaystyle \prod p(\kappa)\displaystyle \prod p(\Gamma)
\end{aligned}
$$

* The number of possible topic structures is exponentially large --> sum is intractable to compute.
* Instead of obtaining a closed-form solution for the posterior distribution, we must approximate it.

@@

* Mean field variational methods (Blei et al., 2001, 2003)
* Expectation propagation (Minka and Lafferty, 2002)
* Collapsed Gibbs sampling (Griffiths and Steyvers, 2002)
* Distributed samplung (Newsman et al., 2008; Ahmed et al., 2012)
* Stochastic inference (Hoffman et al., 2010, 2013; Mimno et al., 2012)
* Factorization inference (Arora et al., 2012; Anandkumar et al., 2012)
* **Variational EM algorithm** (Wang and Blei 2013; Roberts et. al. (2016)
* ...

@@

## Model Selection

**Prior Specification:**

$\gamma_{p,k}$ ~ Normal$(0,\sigma^2_k)$

$\sigma^2_k$ ~ Inverse-Gamma$(a,b)$

<aside class="notes">
The prior specification for the topic prevalence parameters is a zero mena Haussian distribution with shared variance parameter.
$p$ indexes the covariates, $k$ indexes the topics and $a,b$ are fixed hyperparameters
</aside>

**Topic Selection:**

$K=28$

<aside class="notes">
(2)Two potential goals:
(2a)Predict text well. Statistical criteria to select $K$.
(2b)Interpretability. General versus specific.
</aside>

@@

```{r echo=FALSE, fig.height=6.5}
plot(stmOut, type = "summary", xlim = c(0,.16), n = 5, 
     custom.labels = topicLabel$topic_name, 
     family = "Roboto", main = "", text.cex = 1)
```

@@

2. **Relationship between topics** How topics are correlated differently for different newswires, indicating how topics are connected and framed differently in each newswire.

>>>>

---
title: "What is reported in the news?"
author: Franziska Löw
date: December 12, 2017
output: 
  revealjs::revealjs_presentation:
    theme: serif
    highlight: haddock
    center: true
    transition: slide
    css: style.css
    self_contained: true
    incremental: true
---

@@

###1. Introduction

###2. Data

###3. Methodology

###4. First Results

###5. Whats next?

@@

- **Does the business model have an effect on the editorial content?**

# Data

@@

```{r echo=FALSE, fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
btw %>%
  group_by(date) %>%
  summarise(obs = n()) %>%
  ggplot(aes(date, obs)) +
  geom_line(color="blue") +
  geom_vline(aes(xintercept=as.Date("2017-09-24")),
             linetype = 2, color="grey40") +
  geom_vline(aes(xintercept=as.Date("2017-11-20")),
                 linetype = 2, color="grey40") +
  labs(x="", y="Count") +
  scale_x_date(breaks = date_breaks("1 month"), labels=date_format("%B", tz="CET")) +
  theme(
      #legend.position   = "none",
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      #panel.grid        = element_blank(),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )
```

@@

  * $N =$ collection of words within a document.
  * $D =$ collection of documents within a corpus.
  * observed: word in a document $w_{d,n}$
  * fixed: mixture components (number of topics $K$ & vocabulary)
  * hidden: mixture proportions (per-document topic proportions $\theta_d$ & word-topic distribution $\beta_k$)

@@

$$
\begin{aligned}
p(\beta,\theta,z|w) \propto
\displaystyle \prod_{i=1}^{K}p(\beta_i)\displaystyle \prod_{d=1}^{D}p(\theta_d)(\prod_{n=1}^Np(z_{d,n,}|\theta_d)p(w_{d,n}|\beta_{d,k},z_{d,n}))
\end{aligned}
$$

@@

* $K$: choose the number of topics

@@

* $\theta_d$: for each document $d$, choose a distribution over topics;

@@

* $z_{d,n}|\theta_d$: according to $\theta_d$, assign a topic $z_{d,n}$ for the $n^{th}$ word;

@@

* $w_{d,n}|z_{d,n},\beta,\theta$: choose a term from that topic according to $\beta_k$

@@

* $N$: repead this process for all $n$ word-positions in the document.

* $D$: conduct this process for all $d$ documents in the corpus

@@

@@

$$
p(\eta,z,\kappa,\gamma,\Sigma|w,X,Y)\propto (\displaystyle \prod^D_{d=1}(\eta_d|X_d,\gamma,\Sigma)( \prod^N_{n=1}(z_{n,d}|\theta_d)(w_n|\beta_{d,k=z_{d,n}}))) \prod p(\kappa)\prod p(\Gamma)
$$

* The number of possible topic structures is exponentially large
    -> sum is intractable to compute.
* Instead of obtaining a closed-form solution for the posterior distribution, we must approximate it.

@@

* Mean field variational methods (Blei et al., 2001, 2003)
* Expectation propagation (Minka and Lafferty, 2002)
* Collapsed Gibbs sampling (Griffiths and Steyvers, 2002)
* Distributed samplung (Newsman et al., 2008; Ahmed et al., 2012)
* Stochastic inference (Hoffman et al., 2010, 2013; Mimno et al., 2012)
* Factorization inference (Arora et al., 2012; Anandkumar et al., 2012)
* **Variational EM algorithm** (Wang and Blei 2013; Roberts et. al. (2016))
* ...

@@

@@

```{r echo=FALSE, fig.height=6.5}
plot(stmOut, type = "summary", xlim = c(0,.18), n = 5, 
     custom.labels = topicLabel$topic_name, 
     family = "Roboto", main = "", text.cex = 1)
```

@@

2. **Sentiment analysis**

3. **Relationship between topics** How topics are correlated differently for different newswires, indicating how topics are connected and framed differently in each newswire.

# Thank you! Questions?
